Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference

### 2025-01-24 src/tex/ms.tex
[showyourwork] overleaf sync

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/CHEMPY_TNG_yields_Nstar_comp.pdf}
    \vspace{-.5cm}
    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line). For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
    \label{fig:CHEMPY_TNG_N_star_analysis}
    \script{chempy_tng_inference.py}
\end{figure*}

\paragraph{Inference on the \texttt{CHEMPY} TNG yield set}

@@

\subsection{Alternative yield set}
\label{subsec: mocks_wrong_yield}

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/CHEMPY_TNG_yields_Nstar_comp.pdf}
    \vspace{-.5cm}
    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line). For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
    \label{fig:CHEMPY_TNG_N_star_analysis}
    \script{chempy_tng_inference.py}
\end{figure*}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/CHEMPY_alternative_yields_Nstar_comp.pdf}
    \vspace{-.5cm}
    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line). For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
    \label{fig:CHEMPY_TNG_N_star_analysis}
    \script{chempy_alt_yield_inference.py}
\end{figure*}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/TNG_sim_Nstar_comp.pdf}
    \vspace{-.5cm}
    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line). For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
    \label{fig:TNG_N_star_analysis}
    \script{tng_inference.py}
\end{figure*}

\paragraph{Inference on mock data from \texttt{CHEMPY} with TNG yield set}

@@

\subsection{Inference on mock data from \texttt{CHEMPY} with incorrect yield set}
\label{subsec: mocks_wrong_yield}

\begin{figure}
%\vspace{-1.3cm}
    \centering
    \includegraphics[width=\columnwidth]{figures/CHEMPY_alternative_yields.png}
    \vspace{-.25cm}
    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars. The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
    \label{fig:CHEMPY_TNG_sbi} 
    \script{chempy_alt_yield_inference.py}
\end{figure}

\subsection{Inference on mock data from the IllustrisTNG simulation}
\label{subsec: tng_sim}

\begin{figure}
%\vspace{-1.3cm}
    \centering
    \includegraphics[width=\columnwidth]{figures/TNG_simulation.png}
    \vspace{-.25cm}
    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars. The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
    \label{fig:CHEMPY_TNG_sbi} 
    \script{tng_inference.py}
\end{figure}
